---
title: "Reproducible Research Assignment 1"
output: github_document
---

#### Loading necessary packages for the analysis 
```{r}
  library(dplyr)
  library(ggplot2) 
```

#### Loading and processing the data 

Obtaining summary statistics
```{r}
    activity <- read.csv("activity.csv")
    summary(activity) 
    ```

#### What is mean total number of steps taken per day? 

1. Histogram of mean steps taken per day
```{r}
stepsbyday <- aggregate(activity["steps"], by=activity["date"], sum)
hist(stepsbyday$steps, main = "Daily Steps", xlab = "Total Daily Steps") 
```

2. Mean and median steps taken per day across all days
```{r} 
mean(stepsbyday$steps, na.rm = TRUE)
median(stepsbyday$steps, na.rm = TRUE) 
```

#### What is the average activity pattern? 
Time series plot of the average number of steps taken
```{r} 
meanstepsinterval <- aggregate(activity["steps"], by=activity["interval"], mean, na.rm = TRUE)
  plot(meanstepsinterval$interval, meanstepsinterval$steps, type = "l", xlab = "Interval", 
     ylab = "Avg. Steps Accross All Days", main = "Average Daily Activity Pattern") 
```
     
#### Which 5-minute interval, on average, contains the maximum number of steps? 
```{r} 
  meanstepsinterval$interval[which.max(meanstepsinterval$steps)] 
```

#### Imputing missing values 

1. Copying the original dataframe to be manipulated
```{r} 
activityimpute <- read.csv("activity.csv") 
```

2. Replacing missing values with the avaerage of the corresponding 5-minute interval accross all days
```{r} 
activityimpute$steps[is.na(activityimpute$steps)] <-   meanstepsinterval$steps[match(activityimpute$interval,meanstepsinterval$interval)][which(is.na(activityimpute$steps))] 
```

#### What is the total number of steps taken per day with the imputed dataset? 
```{r}
stepsbyday_imputed <- aggregate(activityimpute["steps"], by=activityimpute["date"], sum)
hist(stepsbyday_imputed$steps, main = "Daily Steps (Imputed Data)", xlab = "Total Daily Steps") 
```

#### Are there differences in activity patterns between weekdays and weekends? 

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
activityimpute$typeday <- ifelse(weekdays(as.Date(activityimpute$date)) == "Saturday" | weekdays(as.Date(activityimpute$date)) == "Sunday", "weekend", "weekday") 
meanstepsdaytype <- aggregate(activityimpute$steps ~ activityimpute$interval + activityimpute$typeday, df = activityimpute, FUN = mean)
meanstepsdaytype <- rename(meanstepsdaytype, steps = "activityimpute$steps")
meanstepsdaytype <- rename(meanstepsdaytype, interval = "activityimpute$interval")
meanstepsdaytype <- rename(meanstepsdaytype, typeday = "activityimpute$typeday")
ggplot(data = meanstepsdaytype, aes(x = interval, y = steps)) +
  geom_line() + labs (title= "Average Daily Activity by Weekday", y= "Avg. Steps", x= "Interval") + facet_wrap(~ typeday) 
```
